---
title: "HW2"
author: "Blain Morin"
date: "11/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)


```

**1. Recall the situation from the first problem set, in which $X_i \sim Poisson(\lambda)$ for $i = 1, 2, 3$ with $X_1 = 1, X_2 = 0$ and $X_3 = 3$.**

**(a) Find (analytically) the posterior distribution if the prior distribution on $\lambda$ is a gamma distribution with a shape parameter of 2 and a rate parameter of 10, using the fact that the gamma distribution is the conjugate prior to the Poisson distribution.**

From Jackman, the posterior density is $Gamma(a^*, b^*)$ where $a^* = a + S$ and $b^* = b + n$. 

$$  S = \sum_{1}^{3} X_i =  1 + 0+3 = 4$$

Thus the posterior density is:
$$  Gamma(2+4, 10+3) = Gamma(6, 13) $$

**(b) Use the previous two results to compute the posterior probability that  $\lambda > 1$ under this prior.**

Here, I draw 100,000 sample from the posterior gamma distribution and calculate the probability that the draw is greater than 1:

```{r, echo = TRUE}

### Set seed for reproducibility
set.seed(1111)

sum(rgamma(100000,
           shape = 6,
           rate = 13) > 1) / 100000

```

Thus, the probability that $\lambda > 1$ is .01. This result is comparable to the first HW.

