---
title: "HW2"
author: "Blain Morin"
date: "11/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)


```

**1. Recall the situation from the first problem set, in which $X_i \sim Poisson(\lambda)$ for $i = 1, 2, 3$ with $X_1 = 1, X_2 = 0$ and $X_3 = 3$.**

**(a) Find (analytically) the posterior distribution if the prior distribution on $\lambda$ is a gamma distribution with a shape parameter of 2 and a rate parameter of 10, using the fact that the gamma distribution is the conjugate prior to the Poisson distribution.**

From Jackman, the posterior density is $Gamma(a^*, b^*)$ where $a^* = a + S$ and $b^* = b + n$. 

$$  S = \sum_{1}^{3} X_i =  1 + 0+3 = 4$$

Thus the posterior density is:
$$  Gamma(2+4, 10+3) = Gamma(6, 13) $$

**(b) Use the previous two results to compute the posterior probability that  $\lambda > 1$ under this prior.**

Here, I draw 100,000 sample from the posterior gamma distribution and calculate the probability that the draw is greater than 1:

```{r, echo = TRUE}

### Set seed for reproducibility
set.seed(1111)

sum(rgamma(100000,
           shape = 6,
           rate = 13) > 1) / 100000

```

Thus, the probability that $\lambda > 1$ is .01. This result is comparable to the first HW.

**2. The Wernstrom distribution is a continuous distribution over the positive real numbers with probability density function:**

$$  f_W(x) = \frac{5}{2}sin(x)^2e^{-x} $$
**for all $x>0$.**

**(a) Find a constant, $c$, such that $ce^{-x} \geq f_W(x)$ for all $x>0$.**

Because $sin(x)^2$ is bound by 0 and 1, $c = \frac{5}{2}$.

**(b) Use this information to draw at least 10,000 samples from the Wernstrom distribution using rejection sampling.**

**i. Plot a kernel density estimate of the distribution of your samples or a histogram of your samples.**

**ii. Add the p.d.f. of the Wernstrom distribution to this plot, which should closely match the distribution of your samples.**


```{r, echo = TRUE}

### Set seed 
set.seed(1212)

X = rexp(100000, 1)
U = rexp(100000, 1)

### Wernstrom Distribution
pi_x <- function(x) {
  new_x = (5/2) * (sin(x)^2) * exp(-x)
  return(new_x)
}

count = 1
accept = c()


while(count <= 100000 & length(accept) < 10000){
  test_u = U[count]
  test_x = pi_x(X[count])/(2.5*dexp(X[count],1))
  if (test_u <= test_x){
    accept = rbind(accept, X[count])
    count = count + 1
  }
  count = count + 1
}

### Code adapted from https://rpubs.com/mathetal/rejectsampling

df = data.frame(accept = accept)

f = function(x) (5/2) * (sin(x)^2) * exp(-x)

df %>%
  ggplot(aes(x = accept)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "white") + 
  stat_function(fun = f, color = "red", size = 1.2) +
  theme_bw() +
  ylab("Density") +
  xlab("x") +
  ggtitle("Histogram: Wernstrom Density using Rejection Sampling (Red = Wernstrom Density)")
  
```

As we see in the histogram above, our rejection sampling is successful in approximating the Wernstrom distribution (shown as the red line).

**(c) Use these samples to estimate the mean and variance of the Wernstrom distribution.**

```{r, echo = TRUE}

mean(df$accept)
var(df$accept)

```

We estimate the mean of the Wernstrom distribution to be 1.37 with a variance of .76.